Article Title

### 2014-07-15 manuscript.tex
submitted

To test this hypothesis, we computed the relative cross-validation loss of estimators  $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ in $n=27$ imaged sites in 14 mice.  The hyperparameters of each estimator were optimized by nested cross-validation (See Fig.~\ref{supp:1} and  Methods). Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:3}). The respective median differences of the cross-validation loss were 0.039, 0.0016, 0.0029, and 0.0059 nats/cell/bin, significantly greater than zero ($p<0.01$ in each comparison, $n=27$ sites in 14 mice, Wilcoxon signed rank test).

The same evaluation based on the quadratic loss function,
\begin{equation}\label{eq:quad}
\loss{C,C_{\sf sample}^\prime}=\frac 1 {p^2}\Tr(C^{-1}C_{\sf sample}^\prime-I)^2,
\end{equation}
instead of the normal loss function reproduced the same relationship between the estimators (Fig.~S\ref{supp:02}). This suggests that the results of this study are robust to the choice of the loss function and do not depend on the assumption of gaussianity.

@@

The average partial correlations calculated from these estimates according to Eq.~\ref{eq:partial} at all 27 sites were about 5 times lower than the average sample correlations (Fig.~\ref{fig:5} C). This suggests that correlations between neurons build up from multiple chains of smaller interactions. Furthermore, the average partial correlations were less variable: the coefficient of variation of the average sample correlations across sites was 0.45 whereas that of the average partial correlations was 0.29, with larger populations exhibiting more similar values of the average partial correlations than the smaller populations.

@@

\subsubsection*{Relationship of $C_{\sf sparse+latent}$ to orientation tuning and physical distances}

@@

\section*{Discussion}

@@

@@

\section*{Acknowledgments}

@@

\section*{Figure Legends}
% This section is for figure legends only, do not include
% graphics in your manuscript file.

\begin{figure}
\caption{
{\bf Regularized estimators whose structure matches the true structure in the data are more efficient.}
     {\bf Row 1.} Graphical models of the target estimates of the four respective regularized covariance matrix estimators.  Recorded neurons are represented by the green spheres and latent units by the lightly shaded spheres.  Edges represent non-zero partial correlations, \emph{i.e.}\;`interactions'.
     {\bf Row 1, A}.  For estimator $C_{\sf diag}$, the target estimate is a diagonal matrix, which describes systems that lack linear dependencies.
     {\bf  Row 1, B.} For estimator $C_{\sf factor}$, the target estimate is a factor model (low-rank matrix plus a diagonal matrix), representing systems in which correlations arise due to common input from latent units.
     {\bf  Row 1, C}. For estimator $C_{\sf sparse}$, the covariance matrix is approximated as the inverse of a sparse matrix. This approximation describes systems in which correlations arise from a sparse set of  linear associations between the observed units.
     {\bf  Row 1, D}.  For estimator $C_{\sf sparse+latent}$, the covariance matrix is approximated as the inverse of the sum of a sparse matrix and a low-rank matrix. This approximation describes a model wherein correlations arise due to sparse associations between the recorded cells \emph{and} due to several latent units.
     \\
     {\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: {\bf A.} the diagonal correlation matrix, {\bf B.} a factor model with four latent units, {\bf C.}  a correlation matrix with 67\%  off-diagonal zeros in its inverse, and {\bf  D.} a correlation matrix whose inverse is the sum of a rank-3 matrix (\emph{i.e.}\;three latent units) and a sparse matrix with 76\% off-diagonal zeros.
     \\
{\bf Row 3:} Sample correlation matrices calculated from samples of size $n=500$ drawn from simulated random processes with respective correlation matrices shown in Row 2.  The structure of the sample correlation matrix is difficult to discern by eye.
     \\
{\bf Row 4:} Estimates computed from the same data as in Row 3 using structured estimators of the correct type, optimized by cross-validation.  The regularized estimates are closer to the truth than the sample correlation matrices.
     \\
{\bf Row 5:} Excess losses (Eq.~\ref{eq:loss}) for the five estimators as a function of sample size. The error bars indicate the standard error of the mean based on 30 samples.  Estimators with structure that matches the true model converged to zero faster than the other estimators.
     \\
{\bf Row 6:} Validation losses for the five estimators relative to the matching estimator. Error bars indicate the standard error of the mean based on 30 samples.  Differences in validation loss approximate differences in true loss.
}
\label{fig:1}
\end{figure}

>>>>

To test this hypothesis, we computed the relative cross-validation loss of estimators  $C_{\sf sample}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ in $n=27$ imaged sites in 14 mice.  The hyperparameters of each estimator were optimized by nested cross-validation (See Fig.~S1 and  Methods). Indeed, the sparse+latent estimator outperformed the other estimators (Fig.~\ref{fig:3}). The respective median differences of the cross-validation loss were 0.039, 0.0016, 0.0029, and 0.0059 nats/cell/bin, significantly greater than zero ($p<0.01$ in each comparison, $n=27$ sites in 14 mice, Wilcoxon signed rank test).

The same evaluation based on the quadratic loss function,
\begin{equation}\label{eq:quad}
\loss{C,C_{\sf sample}^\prime}=\frac 1 {p^2}\Tr(C^{-1}C_{\sf sample}^\prime-I)^2,
\end{equation}
instead of the normal loss function reproduced the same relationship between the estimators (Fig.~S2). This suggests that the results of this study are robust to the choice of the loss function and do not depend on the assumption of gaussianity.

@@

The average partial correlations calculated from these estimates according to Eq.~\ref{eq:partial} at all 27 sites were about 5 times lower than the average sample correlations (Fig.~\ref{fig:5} C). This suggests that correlations between neurons build up from multiple chains of smaller interactions. Furthermore, the average partial correlations were less variable: the coefficient of variation of the average sample correlations across sites was 0.45 whereas that of the average partial correlations was 0.29, with larger populations exhibiting greater uniformity of average partial correlations than the smaller populations.

@@

\subsection*{Relationship of $C_{\sf sparse+latent}$ to orientation tuning and physical distances}

@@

\section*{Discussion}
\subsection*{Functional connectivity as a network of pairwise interactions}
Functional connectivity is often represented as a graph of pairwise interactions. The goal of many studies of functional connectivity has been to estimate  anatomical connectivity from  observed multineuronal spiking activity.  For example, characteristic peaks and troughs in the pairwise cross-correlograms of recorded spike trains contain statistical signatures of directional monosynaptic connections and shared synaptic inputs \cite{Gerstein:1964, Perkel:1967, Moore:1970, Alonso:1998, Denman:2013}.  Such signatures are ambiguous as they can arise from network effects other than direct synaptic connections \cite{Aertsen:1989}.  With simultaneous recordings from more neurons, ambiguities can be resolved by inferring the conditional dependencies between pairs of neurons.  Direct causal interactions between neurons produce statistical dependency between them even after conditioning on the state of the remainder of the network and external input. Therefore, conditional independence can signify the absence of a direct causal influence.

Conditional dependencies can be inferred by fitting a probabilistic model of the joint population activity. For example, generalized linear models (GLMs) have been constructed to  include biophysically plausible synaptic integration, membrane kinetics, and individual neurons' stimulus drive~\cite{Pillow:2008}.  Maximum entropy models constrained by observed pairwise correlations are among other models with pairwise coupling between cells \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009}.  Assuming that the population response follows a multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.   Each probabilistic model, fitted to the same data may reveal a completely different network of `interactions',  \emph{i.e.}\;conditional dependencies between pairs of cells.

It is not yet clear which approach provides the best correspondence with anatomical connectivity. Little experimental evidence is available to answer this question.  The connectivity graphs inferred by various statistical methods are commonly reported without examining their relation to anatomy.
Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013, Sadovsky:2014}.  However, the topological properties of functional connectivity graphs can depend on the method of inference \cite{Zalesky:2012}. Until a physiological interpretation of functional connectivity is established, the physiological relevance of such analyses remains in question and we did not attempt them here.

Inference of the conditional dependencies also depends on the completeness of the recorded population:  To equate conditional dependencies to direct interaction between neurons, we must record from all neurons with which the pair interacts. Unobserved portions of the circuit may manifest as conditional dependencies between observed neurons that do not interact. For this reason, statistical models of population activity have been most successfully applied to \emph{in vitro} preparations of the retina or cell cultures where high-quality recordings from the complete populations were available \cite{Pillow:2008}. In cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of inference of the pairwise conditional dependencies. Perhaps for this reason, partial correlations have not, until now, been used to describe the functional connectivity in cortical populations.

Two-photon imaging of population calcium signals presents unique advantages for the estimation of functional connectivity.  While the temporal resolution of calcium signals is limited by calcium dye kinetics, fast imaging techniques combined with spike inference algorithms provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010}. However, such high temporal precision comes at the cost of the accuracy of inferred spike rates.  Better accuracy is achieved when calcium signals are analyzed on scales of tens of milliseconds \cite{Cotton:2013}.  The major advantage of calcium imaging is its ability to characterize the spatial arrangement and types of recorded cells.  Recently, advanced imaging techniques have allowed recording from nearly every cell in a volume of cortical tissue  \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.  These techniques may provide more incisive measurements of functional connectivity than electrophysiological recordings.

The low temporal resolution of calcium signals limits the use of functional connectivity methods that rely on millisecond-scale binning of signals (cross-correlograms, some GLMs, and binary maximum entropy models).  Hence, most studies of functional connectivity have relied on instantaneous sample correlations \cite{Greenberg:2008, Golshani:2009, Hofer:2011, Malmersjo:2013} .  Although some investigators have interpreted such correlations as indicators of (chemical or electrical) synaptic connectivity, most used them as more general indicators of functional connectivity without relating them to underlying mechanisms.

In this study, we sought to infer pairwise functional connectivity networks  in cortical microcircuits. We hypothesized that partial correlations correspond more closely to underlying mechanisms than sample correlations when recordings are sufficiently dense.  Since neurons form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we \emph{a priori} favored solutions with sparse partial correlations.  Under the assumptions that the recorded population is sufficiently complete and that the model correctly represents the nature of interactions, the network of partial correlations can be hypothesized to be a better representation of functional dependencies than correlations.

\subsection*{Functional connectivity as coactivations}
Another approach to describing the functional connectivity of a circuit is to isolate individual patterns of multineuronal coactivations \cite{Gerstein:1989, Chapin:1999, Peyrache:2010, Ch:2010, Lopes:2011, Lopes:2013}. Depending on the method of their extraction, coactivation patterns may be referred to as \emph{assemblies}, \emph{factor loadings}, \emph{principal components}, \emph{independent components}, \emph{activity modes}, \emph{eigenvectors}, or \emph{coactivation maps}. Coactivation patterns could be interpreted as signatures of Hebbian cell assemblies \cite{Gerstein:1989, Ch:2010}, \emph{i.e.}\;groups of tightly interconnected groups of cells involved in a common computation.  Coactivation patterns could also result from shared input from unobserved parts of the circuit, or global network fluctuations modulating the activity of the local circuit \cite{Okun:2012}.

Coactivation patterns and pairwise connectivity are not mutually exclusive since assemblies arise from patterns of synaptic connectivity.  However, an analysis of coactivation shifts the focus from detailed interactions to  collective behavior.
In our study, the functional connectivity solely through modes of coactivations was represented by the factor analysis-based estimator $C_{\sf factor}$.

\subsection*{Combining pairwise interactions and coactivations}
In the effort to account for the joint activity patterns that are poorly explained by pairwise interactions, investigators have augmented models of pairwise interactions with additional mechanisms such as latent variables \cite{Koster:2013},  high-order correlations \cite{Ganmor:2011}, or global network fluctuations \cite{Tkacik:2013}.

In our study, we combined pairwise interactions with collective coactivations by applying the recently developed numerical techniques for the inference of the partial correlation structure in systems with latent variables \cite{Chandrasekaran:2010, Ma:2013}.  The resulting estimator, $C_{\sf sparse+latent}$, effectively decomposed the functional connectivity into a sparse network of pairwise interactions and coactivation mode vectors.

\subsection*{Addressing ill-posedness}
Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data produce large errors in the inferred network of dependencies. The problem becomes worse as the number of  recorded neurons increases until such models lose their statistical validity \cite{Roudi:2009}.  As techniques have improved to allow recording from larger neuronal populations, experimental neuroscientists have addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted \cite{Tkacik:2013}. However, ambitious projects, such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations. Simply increasing recording duration will not be sufficient, and the problem must be addressed by using regularized estimators. Regularization biases the solution toward a small subspace in order to counteract the effect of  sampling noise in the empirical data. However, biasing the solution to an inappropriate subspace does not allow significant estimation improvement and hinders interpretation.

Several strategies have been developed to limit the model space in order to improve the quality of the estimate. For example, Ganmor et al. \cite{Ganmor:2011} developed a heuristic rule to identify the most significant features that must be fitted by a maximum entropy model for improved performance in the retina. As another example of regularization, generalized linear models typically employ $L_1$ penalty terms to constrain the solution space and to effectively reduce the dimensionality of the solution \cite{Pillow:2008}.

In our study, regularizations were accomplished by dimensionality reduction (feature selection) schemes to produce sparse, constrained solutions. Only the most efficient scheme was considered in the analysis of functional connectivity.

\subsection*{Model selection}
Various model selection criteria have been devised to select between families of models and the optimal subsets of variables in a given model family based on observed data. Despite its computational requirements, cross-validation is among the most popular model selection methods due to its minimal assumptions about the data generating process \cite{Arlot:2010}.

We evaluated the covariance matrix estimators a loss function based on multivariate normal log likelihood.  However, this does not limit the applicability of its conclusions to normal distributions. Indeed our major findings could be reproduced using other loss functions (compare Fig.~\ref{fig:3} and Fig.~S2).  Other probabilistic models, fitted to the same data, could also serve as estimators of the covariance matrix.  If a different model yields better estimation of the covariance matrix than the estimator proposed here, we believe that its structure should deserve consideration as the better representation of the functional connectivity.

The results of model selection must be interpreted with caution.  As we demonstrated by simulation, even models with incorrect forms of dependencies can substantially improve estimates (Fig.~\ref{fig:1}). Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not necessarily support the conclusion that it reveals a better representation of dependencies in the data.  This caveat is related to \emph{Stein's Paradox} \cite{Efron:1977}: The biasing of an estimate toward an arbitrary low-dimensional target can consistently outperform a less constrained estimate.

\subsection*{Physiological interpretation and future directions}

We showed that among several models a sparse network of linear interactions with several latent inputs yielded the best estimates of the noise covariance matrix for cortical microcircuits.  This finding is valuable in itself: improved estimates of the noise covariance matrix for large datasets are important in order to understand the role of noise correlations in population coding \cite{Abbott:1999, Sompolinsky:2001, Averbeck:2006, Ecker:2011}

Moreover, this estimation approach provides a graphical representation of the dependencies in the data that can be used to formulate and test hypotheses about the structure of connectivity in the microcircuit. Importantly, the inferred functional interactions were substantially different from the network of the highest sample correlations.  For example, the $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that were not present in the sample correlation matrix (Fig.~\ref{fig:4} F) and may reflect inhibitory circuitry.

Distances between cells in physical space and in sensory feature space had a stronger effect on the partial correlations estimated by the $C_{\sf sparse+latent}$ estimator than on sample correlations (Fig.~\ref{fig:6} A--C).
These differences support the idea that correlations are built up from partial correlations in chains of intermediate cells positioned closer and tuned more similarly to one another, with potentially closer correspondence to anatomical connectivity.  These differences may also be at least partially explained by a trivial effect of regularization: the $L_1$ penalty applied by the estimator (Eq.~\ref{eq:ma}) suppresses small partial correlations to greater extent than large partial correlations, enhancing the apparent effect of distance and tuning.  Distances between cells in physical space and in sensory feature space also had distinct effects on the positive and negative connectivities inferred by the estimator (Fig.~\ref{fig:6} D--F); we could not think of a numerical artifact that could produce this effect. The distinct positive and negative connectivity patterns  may reflect geometric and graphical features of local excitatory and inhibitory networks. Indeed, the relationships between patterns of positive and negative connectivities inferred by the estimator mirrored the properties of excitatory and inhibitory synaptic connectivities with respect to distance, cortical layers, and feature tuning \cite{Song:2005, Oswald:2008, Adesnik:2010, Perin:2011, Fino:2011, Hofer:2011, Isaacson:2011, Levy:2012}. For example, while excitatory neurons form synapses within highly specific local cliques \cite{Perin:2011}, inhibitory interneurons form synapses with nearly all excitatory cells within local microcircuits \cite{Fino:2011, Hofer:2011, Packer:2011}.  To further investigate the link between synaptic connectivity and inferred functional connectivity, in future experiments, we will use molecular markers for various cell types with follow-up multiple whole-cell \emph{in vitro} recordings \cite{Hofer:2011, Ko:2013} to directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry. Finally, the latent units inferred by the estimator can be analyzed for their physiological functions. For example, these latent units may be modulated under different brain states (e.g. slow-wave sleep, attention) and stimulus conditions (e.g. certain types of stimuli may engage feedback connections).

@@

\subsection*{Ethics statement}
All procedures were conducted in accordance with the ethical guidelines of the National Institutes of Health and were approved by the Baylor College of Medicine IACUC.

\subsection*{Surgery and two-photon imaging}
The surgical procedures and data acquisition were performed as described in \cite{Cotton:2013}: C57BL/6J mice (aged p40--60) were used. Anesthesia was initiated with isoflurane (3\%) and maintained with the mixture of fentanyl (0.05 mg/kg), midazolam (5 mg/kg), and medetomidine (0.5 mg/kg), with boosts of half the initial dose every 3 hours.  A craniotomy was performed over the right primary visual cortex.  Membrane-permeant calcium indicator Oregon Green 488 BAPTA-1 AM (OGB-1, Invitrogen) was loaded by bolus injection.  The craniotomy was sealed using a glass coverslip secured with dental cement.

Calcium imaging began 1 hour after dye injection.  All imaging was performed using 3D-RAMP two-photon microscopy \cite{Cotton:2013}. First, a 3D stack was acquired and cells were manually segmented. Then calcium signal were collected by sampling in the center of each cell at rates of 100 Hz or higher, depending on the number of cells.

\subsection*{Visual stimulus}
The visual stimulus consisted of full-field drifting gratings with 90\% contrast, luminance of 10 cd/m$^2$, spatial frequency of 0.08 cycles/degree, and temporal frequency of 2 cycles/s. Two types of stimuli were presented for each imaging site: First, directional tuning was mapped using a pseudo-random sequence of drifting gratings at sixteen directions of motion, 500 ms per direction, without blanks, with 12--30 trials for each direction of motion.  Second, ro measure correlations, the stimulus was modified to include only two directions of motion (in 9 datasets) or five directions (in 22 datasets) and the gratings were presented for 1 second and were separated by 1-second blanks, with 100--300 trials for each direction of motion.

\subsection*{Data processing}
All data were processed in MATLAB using the DataJoint data processing chain toolbox (http://datajoint.github.com).

The measured fluorescent traces were deconvolved to reconstruct the firing rates for each neuron: First, the first principal component was subtracted from the raw traces in order to reduce common mode noise related to small cardiovascular movements \cite{Cotton:2013}. The resulting traces were high-pass filtered above 0.1 Hz and downsampled to 20 Hz (Fig.~\ref{fig:2} C). Then, the firing rates were estimated using by nonnegative deconvolution \cite{Vogelstein:2010}.

Orientation tuning was computed by fitting the mean firing rates for each direction of motion $\phi$ using two-peaked von Mises tuning functions $f(\phi)=a + b\exp\left[\frac 1 w(\cos(\phi-\theta)-1) \right] + c\exp\left[\frac 1 w(\cos(\phi-\theta+\pi)-1) \right]$ where $b\ge c$ are the amplitudes of the two respective peaks, $w$ is the tuning width, and  $\theta$ is the preferred direction. The significance of the fit was determined by the permutation test: the labels of the direction were randomly permuted 10,000 times; the $p$-values of the fits were computed as the fraction of permutations that yielded $R^2$ equal to or higher than that of the original data.  Cells were considered tuned with $p<0.05$.

For covariance estimation, the analysis was limited to the period with 2 or 5 stimulus conditions and lasted between 14 and 27 minutes (mean 22 minutes).  Cells that did not have substantial spiking activity (those whose variance was less than 1\% of the median across the site) or whose activity was unstable (those whose variance in the least active quarter of the recording did not exceed 1\% of the variance in the most active quarter) were excluded from the analysis.

\subsection*{Cross-validation}
To compare the performance of the estimators, we used conventional 10-fold cross-validation: Trials were randomly divided into 10 subsets with approximately equal numbers of trials of each condition in each subset. Each subset was then used as the testing sample with the rest of the data used as the training sample for estimating the covariance matrix. The average validation loss over the 10 folds was reported.

Since each of the regularized estimators had one or two hyperparameters, we used \emph{nested cross-validation}:  The outer loop evaluated the performance of the estimators with the hyperparameter values optimized by cross-validation within the inner loop.  Hyperparameters were optimized by a two-phase search algorithm: random search to find a good starting point for the subsequent pattern search to find the global minimum.  The inner cross-validation loop subdivided the training dataset from the outer loop to perform 10-fold cross-validation in order to evaluate each choice of the hyperparameter values.  Thus the size of the training dataset within the inner loop comprised 81\% of the entire recording. Fig.~S1 illustrates the dependence of the validation loss on the hyperparameters of the $C_{\sf sparse+latent}$ estimator for the example site shown in Figures \ref{fig:2} and \ref{fig:4} and the optimal value found by the pattern search algorithm.

When the validation loss was not required, only the inner loop of cross-validation was used on the entire dataset.  This approach was used to compute the covariance matrix estimates and their excess loss in the simulation study (Fig.~\ref{fig:1} Rows 4 and 5) and to analyze the partial correlation structure of the $C_{\sf sparse+latent}$ estimator (Fig.~\ref{fig:4}--\ref{fig:6}).

\subsection*{Covariance estimation}
Within the inner loop of cross-validation, regularized covariance matrix estimation required only the sample covariance matrix $C_{\sf sample}$ of the training dataset and the hyperparameter values provided by the outer loop.

Estimator $C_{\sf diag}$ (Eq.~\ref{eq:c-diag})  used two hyperparameters: the covariance shrinkage intensity $\lambda \in [0,1]$ and variance shrinkage intensity $\alpha \in [0,1]$.  The variances (the diagonal of $C_{\sf sample}$) were shrunk linearly toward their mean value $\frac 1 p \Tr(C_{\sf sample})$:
\begin{equation}
D = (1-\alpha)C_{\sf sample}\circ I + \alpha \frac 1 p \Tr(C_{\sf sample}) I
\end{equation}
The $C_{\sf diag}$ estimate was then obtained by shrinking $C_{\sf sample}$ toward $D$ according to Eq.~\ref{eq:c-diag}.

In estimator $C_{\sf factor}$ (Eq.~\ref{eq:c-factor}), the low-rank matrix $L$ and the diagonal matrix $D$ are found by solving the minimization problem
\begin{equation}
(L,D) = \argmin\limits_{\hat L,\hat D} \loss{\hat L + \hat D,C_{\sf sample}},
\end{equation}
by an expectation-maximization (EM) algorithm with specified rank of $L$. Additional the diagonal matrix of individual variances is shrunk toward its mean value according to Eq.~\ref{eq:c-factor}.

In estimator $C_{\sf sparse}$ (Eq.~\ref{eq:c-sparse}), the sparse precision matrix $S$ is found by minimizing the $L_1$-penalized loss with regularization parameter $\lambda$:
\begin{equation}
S = \argmin\limits_{\hat S \succ 0} \loss{{\hat S}^{-1},C_{\sf sample}} + \lambda \|\hat S \|_1
\end{equation}
where $\hat S\succ 0$ denotes the constraint that $\hat S$ be a positive-definite matrix and $\|\hat S\|_1$ is the element-wise $L_1$ norm of the matrix $\hat S$. This problem formulation is known as \emph{graphical lasso} \cite{Meinshausen:2006, Friedman:2008}. To solve this minimization problem, we adapted the alternative-direction method of multipliers (ADMM) \cite{Ma:2013}.
Unlike $C_{\sf diag}$ and $C_{\sf factor}$, this estimator does not include linear shrinkage: the selection of the sparsity level provides sufficient flexibility to fine-tune the regularization level.

Estimator $C_{\sf sparse+latent}$ (Eq.~\ref{eq:c-sl}) estimates a larger sparse precision matrix $S^\ast$ of the joint distribution of the $p$ observed neurons and $d$ latent units.
\begin{equation}
S^\ast=
\begin{pmatrix}
S & S_{12} \\
S_{12}^\T & S_{22}
\end{pmatrix},
\end{equation}
where the $p\times p$ partition $S$ corresponds to the visible units.
Then the covariance matrix of the observed population is
\begin{equation}
C_{\sf sparse+latent} = \left(S-S_{12}S_{22}^{-1}S_{12}^\T\right)^{-1}
\end{equation}
The rank of the $p\times p$  matrix $L=S_{12}S_{22}^{-1}S_{12}^\T$ matches the number of the latent units in the joint distribution. Rather than finding $S_{12}$ and $S_{22}$ separately, $L$ can be estimated as a low-rank matrix. We adapted the ADMM algorithm to minimize the loss function with $L_1$ penalty on $S$ to regulate its sparseness and nuclear norm penalty on $L$ to regulate its rank \cite{Chandrasekaran:2010,Ma:2013}:
\begin{equation}\label{eq:ma}
(S,L) = \argmin\limits_{\hat S,\hat L} \loss{\hat S-\hat L, C_{\sf sample}} + \alpha\|\hat S\|_1 + \beta\Tr(\hat L)
\end{equation}

The partial correlation matrix (Eq.~\ref{eq:partial}) computed from $C_{\sf sparse+latent}$ includes interactions between the visible and latent units and was used in Fig.~\ref{fig:4} C and D and Fig.~\ref{fig:5} C, and Fig.~\ref{fig:6} A--C).  The partial correlation matrix computed from $S$ alone expresses strengths of pairwise interactions
\begin{equation}
P_{\sf sparse} = -(S\circ I)^{-\frac 1 2} S  (S\circ I)^{-\frac 1 2}
\end{equation}
and were used in Fig.~\ref{fig:4} F, G, H.

The MATLAB code for these computations is available online at http://github.com/atlab/cov-est.
\subsection*{Cross-validation with conditioned variances}
This study required an unbiased estimate of the discrepancy between various correlation matrix estimates and the true value of the correlation matrix. This was accomplished by cross-validation whereby regularized covariance matrix estimates computed from training samples were compared to unbiased covariance matrix estimates computed from independent testing samples. Loss functions were defined with respect to the covariance matrix rather than the correlation matrix because no unbiased estimate of the correlation matrix exists for finite sample sizes in the general case \cite{Fisher:1921}. In particular, we showed that with the normal loss function (Eq.~\ref{eq:loss}), the validation loss $\loss{C,C_{\sf sample}^\prime}$ is an unbiased estimate of true loss when $C_{\sf sample}^\prime$ is an unbiased estimate. However, this approach only worked under the assumption of uniform variances across all conditions included in the training and testing samples.  Here, we extend the validation loss to the estimation and evaluation of a common correlation matrix across multiple conditions with different variances.

Let $T_c$ and $T_c^\prime$ denote the time bin indices for the training and testing samples, respectively, limited to condition $c$. Here, the prime symbol marks quantities estimated from the testing sample.

Similar to Eq.~\ref{eq:sample}, the training and testing sample covariance matrices for condition $c$ are
\begin{equation}
    C_{c,{\sf sample}}
    = \frac 1{n_c} \sum\limits_{t\in T_c}\left(x(t)-\bar x_c\right)\left(x(t)-\bar x_c\right)^\T
\end{equation}
and
\begin{equation}
    C_{c,{\sf sample}}^\prime
    = \frac 1{n_c^\prime} \sum\limits_{t\in T_c^\prime}\left(x(t)-\bar x_c\right)\left(x(t)-\bar x_c\right)^\T
\end{equation}
Here $n_c$ and $n_c^\prime$ denote the sizes of $T_c$ and $T_c^\prime$, respectively.

Note that $\bar x_c= \frac 1 {n_c} \sum\limits_{t \in T_c}x(t)$ is estimated from the training sample but used in both estimates, making $C_{c,\sf sample}^\prime$ an unbiased estimate of the true covariance matrix, $\Sigma$. As such, $C_{c,\sf sample}^\prime$ can be used for validation.

The common correlation matrix $R_{\sf sample}$ is estimated by averaging the condition-specific correlations:
\begin{equation}
    R_{\sf sample}  
    = \frac 1 n \sum\limits_c n_c \left(V_{c,\sf sample}^{-\frac 1 2}C_{c,\sf sample}V_{c,\sf sample}^{-\frac 1 2}\right)
    = \frac 1 n \sum\limits_c \sum\limits_{t \in T_c} z(t)z(t)^\T,
\end{equation}
where $n=\sum\limits_c n_c$ and $V_{c,\sf sample} = C_{c,\sf sample} \circ I$ is the diagonal matrix containing the sample variances. Then $R_{\sf sample}$ is simply the covariance matrix of the $z$-score signal $z(t) = V_{c,\sf sample}^{-\frac 1 2} \left(x(t) - \bar x_c\right)$ of the training sample.

For consistency with prior work, we applied regularization to covariance matrices rather than to correlation matrices. The common covariance matrix was estimated by scaling $R_{\sf sample}$ by the average variances across conditions $V_{\sf sample} = \frac 1 n \sum\limits_c n_c V_{c,\sf sample}$:
\begin{equation}
    C_{\sf sample} = V_{\sf sample}^{\frac 1 2} R_{\sf sample} V_{\sf sample}^{\frac 1 2}
\end{equation}
Note that $C_{\sf sample}$ differs from the sample covariance matrix computed without conditioning the variances on $c$ and this computation helps avoid any biases that would be introduced by ignoring changes in variance.

The covariance matrix estimators $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$ or $C_{\sf sparse+latent}$ convert $C_{\sf sample}$ into its regularized counterpart denoted here as $C_{\sf reg}$.

To evaluate the estimators, we regularized the conditioned variances by linear shrinkage toward their mean value across all conditions. This was done by scaling $C_{\sf reg}$ by the conditioned variance adjustment matrix $Q_c = \delta I + (1-\delta)V_{\sf sample}^{-1} V_{c,\sf sample}$ to produce the conditioned regularized covariance matrix estimate:
\begin{equation}
    C_{c,\sf reg} = Q_c^{\frac 1 2} C_{\sf reg} Q_c^{\frac 1 2}
\end{equation}

The variance regularization parameter $\delta \in [0,1]$ was optimized in the inner loop of cross-validation along with the other hyperparameters.

The overall loss function is obtained by averaging the loss functions across all the conditions:
\begin{equation}\label{eq:full-loss}
    \frac 1 {\sum\limits_c n_c^\prime}\sum\limits_c n_c^\prime \loss{C_{c,\sf reg}, C_{c,\sf sample}^\prime} 
\end{equation}

With the normal loss $\loss{\cdot,\cdot}$ (Eq.~\ref{eq:loss}) and the unbiased testing covariance matrix $C_{c,\sf sample}$, the loss function in Eq.~\ref{eq:full-loss} is an unbiased estimate of the true loss. Hence, it was used for evaluations reported in Fig.~\ref{fig:3}.

\subsection*{Simulation}
For simulation, ground truth covariance matrices were produced by taking 150 independent samples from an artificial population of 50 independent, identically normally distributed units. The covariance matrices were then subjected to the respective regularizations to produce the ground truth matrices for the simulation studies (Fig.~\ref{fig:1} Row 2). Samples were then drawn from multivariate normal distributions with the respective true covariance matrices to be estimated by each of the estimators.

@@

\section*{Acknowledgments}
We thank Genevera Allen for a helpful discussion, and Eftychios Pnevmatikakis for helpful suggestions and feedback on the manuscript.  This work was supported by grants NEI R01-EY018847, NEI P30-EY002520-33, NEI T32-EY07001-37, NIDA R01DA028525, the NIH-Pioneer award DP1-OD008301 to AST, the McKnight Scholar Award to AST; the Arnold \& Beckman Foundation Young Investigator Award to AST; and NSF grants DMS-0817649, DMS-1122094 to KJ.

@@

\newpage
\section*{Figure Legends}
% This section is for figure legends only, do not include
% graphics in your manuscript file.

\begin{figure}[h!]
\caption{
{\bf Regularized estimators whose structure matches the true structure in the data are more efficient.}
     {\bf Row 1.} Graphical models of the target estimates of the four respective regularized covariance matrix estimators.  Recorded neurons are represented by the green spheres and latent units by the lightly shaded spheres.  Edges represent non-zero partial correlations, \emph{i.e.}\;`interactions'.
     {\bf Row 1, A}.  For estimator $C_{\sf diag}$, the target estimate is a diagonal matrix, which describes systems that lack linear dependencies.
     {\bf  Row 1, B.} For estimator $C_{\sf factor}$, the target estimate is a factor model (low-rank matrix plus a diagonal matrix), representing systems in which correlations arise due to common input from latent units.
     {\bf  Row 1, C}. For estimator $C_{\sf sparse}$, the covariance matrix is approximated as the inverse of a sparse matrix. This approximation describes systems in which correlations arise from a sparse set of  linear associations between the observed units.
     {\bf  Row 1, D}.  For estimator $C_{\sf sparse+latent}$, the covariance matrix is approximated as the inverse of the sum of a sparse matrix and a low-rank matrix. This approximation describes a model wherein correlations arise due to sparse associations between the recorded cells \emph{and} due to several latent units.
     \\
     {\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: {\bf A.} the diagonal correlation matrix, {\bf B.} a factor model with four latent units, {\bf C.}  a correlation matrix with 67\%  off-diagonal zeros in its inverse, and {\bf  D.} a correlation matrix whose inverse is the sum of a rank-3 matrix (\emph{i.e.}\;three latent units) and a sparse matrix with 76\% off-diagonal zeros.
     \\
{\bf Row 3:} Sample correlation matrices calculated from samples of size $n=500$ drawn from simulated random processes with respective correlation matrices shown in Row 2.  The structure of the sample correlation matrix is difficult to discern by eye.
     \\
{\bf Row 4:} Estimates computed from the same data as in Row 3 using structured estimators of the correct type, optimized by cross-validation.  The regularized estimates are closer to the truth than the sample correlation matrices.
     \\
{\bf Row 5:} Excess losses (Eq.~\ref{eq:loss}) for the five estimators as a function of sample size. The error bars indicate the standard error of the mean based on 30 samples.  Estimators with structure that matches the true model converged to zero faster than the other estimators.
     \\
{\bf Row 6:} Validation losses for the five estimators relative to the matching estimator. Error bars indicate the standard error of the mean based on 30 samples.  Differences in validation loss approximate differences in true loss.
}
\label{fig:1}
\end{figure}
